---
title: "ST3131 assignment"
output: html_document
date: "2026-03-29"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

```{r}
setwd("C:\\Users\\mxsng\\OneDrive - National University of Singapore\\Desktop\\Min Xuan\\nus\\y2s2\\st3131\\assignment")
data = read.csv("hdb-resale-July-2020.csv")
```

# Replace features using domain knowledge

```{r}
## map the towns to regions CCR, RCR, OCR

library(dplyr)
unique(data$town)

CCR = c("BUKIT TIMAH", "CENTRAL AREA")

RCR = c("BISHAN", "GEYLANG", "KALLANG/WHAMPOA", "MARINE PARADE", "QUEENSTOWN", "TOA PAYOH")

OCR = c("ANG MO KIO", "BEDOK", "BUKIT BATOK", "BUKIT MERAH", "BUKIT PANJANG", "CHOA CHU KANG", "CLEMENTI", "HOUGANG", "JURONG EAST", "JURONG WEST", "PASIR RIS", "PUNGGOL", "SEMBAWANG", "SENGKANG", "SERANGOON", "TAMPINES", "WOODLANDS", "YISHUN")

data <- data %>%
  mutate(region = case_when(
    town %in% CCR ~ "CCR",
    town %in% RCR ~ "RCR",
    town %in% OCR ~ "OCR"
  ))

unique(data$region)
```

```{r}
## map the remaining lease to flat age

library(stringr)

data <- data %>%
  mutate(
    y = as.numeric(str_extract(remaining_lease, "\\d+(?=\\s*years?)")), 
    m = as.numeric(str_extract(remaining_lease, "\\d+(?=\\s*months?)")),
    
    decimal_remaining = coalesce(y, 0) + (coalesce(m, 0) / 12),
    
    flat_age = 99 - decimal_remaining
  ) %>%
  mutate(flat_age = round(flat_age, 2))
```

```{r}
# Group storey_range together
unique(data$storey_range)

low = c("01 TO 03", "04 TO 06", "07 TO 09", "10 TO 12", "13 TO 15")

mid = c("16 TO 18", "19 TO 21", "22 TO 24", "25 TO 27", "28 TO 30")

high = c("31 TO 33", "34 TO 36", "37 TO 39", "40 TO 42", "49 TO 51")

data <- data %>%
  mutate(storey = case_when(
    storey_range %in% low ~ "low",
    storey_range %in% mid ~ "mid",
    storey_range %in% high ~ "high"
  ))

unique(data$storey)
```

```{r}
# Group flat_models
unique(data$flat_model)

standard = c("Standard", "Improved", "New Generation", "Model A", "Simplified", "Model A2")

premium = c("Premium Apartment", "Premium Apartment Loft", "DBSS")

large = c("Maisonette", "Model A-Maisonette", "Apartment", "Adjoined flat", "Multi Generation")

rare = c("Terrace", "Type S1", "Type S2")

data <- data %>%
  mutate(flat_model_group = case_when(
    flat_model %in% standard ~ "standard",
    flat_model %in% premium ~ "premium",
    flat_model %in% large ~ "large",
    flat_model %in% rare ~ "rare"
  ))

unique(data$flat_model_group)
```

```{r}
## Order for storey_range
data$storey <- factor(data$storey, levels = c("low", "mid", "high"))

## Order for flat_type
flat_type_order <- c("1 ROOM", "2 ROOM", "3 ROOM", "4 ROOM", "5 ROOM", "EXECUTIVE", "MULTI-GENERATION")
data$flat_type <- factor(data$flat_type, levels = flat_type_order)

## Order for region
data$region <- factor(data$region, levels = c("OCR", "RCR", "CCR"))

## Order flat_model_group as factor
data$flat_model_group <- factor(data$flat_model_group, levels = c("standard", "premium", "large", "rare"))
```

```{r}
length(unique(data$block)) ##1187 unique blocks for dataset with 2456 obs
## merge the blocks into groups with the same hundreds place to have fewer unique values
data$block_group = floor(as.numeric(gsub("[a-zA-Z]+$", "", data$block)) / 100)
sort(unique(data$block_group))

data$block_group = factor(data$block_group)
```

```{r}
attach(data)
```

# Part 1: EDA

## summary statistics and visualising relationships between resale_price and regressors

```{r}
summary(resale_price)
sd(resale_price)
hist(resale_price)
cor(resale_price, floor_area_sqm)
cor(resale_price, flat_age)

plot(floor_area_sqm, resale_price, main="Scatterplot of resale_price against floor_area_sqm")
abline(lm(resale_price~floor_area_sqm), col="red", lwd = 2)

plot(flat_age, resale_price, main="Scatterplot of resale_price against flat_age")
abline(lm(resale_price~flat_age), col="red", lwd = 2)

boxplot(resale_price~region, main="Boxplot of resale_price~region")

boxplot(resale_price~block_group, main="Boxplot of resale_price~block_group")

par(mar = c(10, 4, 4, 2))
boxplot(resale_price~flat_type, main="Boxplot of resale_price~flat_type", xlab="", las=3)

boxplot(resale_price~storey, main="Boxplot of resale_price~storey")

par(mar = c(10, 4, 4, 2))
boxplot(resale_price~flat_model_group, main="Boxplot of resale_price~flat_model_group")
```

## boxcox to suggest transformation for resale_price

```{r}
library(MASS)
B = boxcox(lm(resale_price~region+block_group+flat_type+storey+floor_area_sqm+flat_model_group+flat_age, data), lambda=seq(-2, 2, 0.5))
B$x[which(B$y == max(B$y))]
```

## visualising relationships between log(resale_price) and regressors after transforming resale_price to log(resale_price)

```{r}
hist(log(resale_price))
cor(log(resale_price), floor_area_sqm)
cor(log(resale_price), flat_age)

plot(floor_area_sqm, log(resale_price), main="Scatterplot of log(resale_price) against floor_area_sqm")
abline(lm(log(resale_price)~floor_area_sqm), col="red", lwd = 2)

plot(flat_age, log(resale_price), main="Scatterplot of log(resale_price) against flat_age")
abline(lm(log(resale_price)~flat_age), col="red", lwd = 2)

boxplot(log(resale_price)~region, main="Boxplot of log(resale_price)~region")

boxplot(log(resale_price)~block_group, main="Boxplot of log(resale_price)~block_group")

par(mar = c(10, 4, 4, 2))
boxplot(log(resale_price)~flat_type, main="Boxplot of log(resale_price)~flat_type", xlab="", las=3)

boxplot(log(resale_price)~storey, main="Boxplot of log(resale_price)~storey")

par(mar = c(10, 4, 4, 2))
boxplot(log(resale_price)~flat_model_group, main="Boxplot of log(resale_price)~flat_model_group")
```

# Part 2 (initial model)

```{r}

model1 = lm(log(resale_price)~region+block_group+flat_type+storey+floor_area_sqm+flat_model_group+flat_age, data)
summary(model1)
anova(model1)
```

## residual analysis for model1

```{r}
sr1 = rstandard(model1)

hist(sr1)
qqnorm(sr1)
qqline(sr1)

plot(floor_area_sqm, sr1, main="Scatterplot of sr1 against floor_area_sqm")
abline(h=0, col="red", lwd = 2)

plot(flat_age, sr1, main="Scatterplot of sr1 against flat_age")
abline(h=0, col="red", lwd = 2)

boxplot(sr1~region, main="Boxplot of sr1~region")
abline(h=0, col="red", lwd = 2)

boxplot(sr1~block_group, main="Boxplot of sr1~block_group")
abline(h=0, col="red", lwd = 2)

par(mar = c(10, 4, 4, 2))
boxplot(sr1~flat_type, main="Boxplot of sr1~flat_type", xlab="", las=3)
abline(h=0, col="red", lwd = 2)

boxplot(sr1~storey, main="Boxplot of sr1~storey")
abline(h=0, col="red", lwd = 2)

par(mar = c(10, 4, 4, 2))
boxplot(sr1~flat_model_group, main="Boxplot of sr1~flat_model_group")
abline(h=0, col="red", lwd = 2)
```

## outlier and influential point detection

```{r}
which(sr1 > 3 | sr1 < -3)
length(which(sr1 > 3 | sr1 < -3))

which(cooks.distance(model1)>1)
length(which(cooks.distance(model1)>1))
```

# Part 2 (final model)

## check multicollinearity using VIF

```{r}
library(car)
vif(model1)
```

## check model adjusted R-squared after dropping flat_type

```{r}
model2 = lm(log(resale_price)~region+block_group+storey+floor_area_sqm+flat_model_group+flat_age, data)
summary(model2)
```

## check multicollinearity after dropping flat_type using VIF

```{r}
vif(model2)
```

## WLS to handle variance heteroscedasticity

```{r}
var_model = lm(abs(residuals(model2))~region+block_group+storey+floor_area_sqm+flat_model_group+flat_age, data)
weights_wls = 1 / (fitted(var_model)^2)

model3 = lm(log(resale_price)~region+block_group+storey+floor_area_sqm+flat_model_group+flat_age, data, weights = weights_wls)
summary(model3)
anova(model3)
```

## residual analysis for model3

```{r}
sr3 = rstandard(model3)

hist(sr3)
qqnorm(sr3)
qqline(sr3)

plot(floor_area_sqm, sr3, main="Scatterplot of sr3 against floor_area_sqm")
abline(h=0, col="red", lwd = 2)

plot(flat_age, sr3, main="Scatterplot of sr3 against flat_age")
abline(h=0, col="red", lwd = 2)

plot(sr3~region, main="Boxplot of sr3~region")
abline(h=0, col="red", lwd = 2)

plot(sr3~block_group, main="Boxplot of sr3~block_group")
abline(h=0, col="red", lwd = 2)

plot(sr3~storey, main="Boxplot of sr3~storey")
abline(h=0, col="red", lwd = 2)

par(mar = c(10, 4, 4, 2))
plot(sr3~flat_model_group, main="Boxplot of sr3~flat_model_group")
abline(h=0, col="red", lwd = 2)
```

## Backward elimination using AIC for variable selection

```{r}
stepAIC(model3, direction="backward")$anova
```
